The Private School Network: Recruiting Visits to Private High Schools by Public and Private Universities

### 2023-11-13 r_r/manuscript_rihe.Rmd
working through revisions

## RQ3: Overlapping network structure

Research question 3 asks, "To what extent do public research universities and selective private universities visit overlapping sets of private high schools?" In contrast to RQ1 and RQ2, answering RQ3 requires social network analysis techniques.

As described in the methods section, we began by constructing a two-mode social network of visits to private high schools (mode 1) by universities (mode 2). Each university and each private high school is a "vertex." A visit from a university to a school is an "edge." Because large two-mode networks like our school-university network are difficult to visualize, two-mode networks are often analyzed as one-mode networks [@RN2457;@RN4663]. We transformed our two-mode school-university network into a one-mode university network in which each university is a vertex and two universities share an edge if both universities visited at least one high school in common. The *weight* of an edge indicates the number of high schools that both universities visited. For example, if there are 200 private high schools that received a visit from both Notre Dame and Villanova, the weight of the edge between Notre Dame and Villanova is 200. We created three separate one-mode university networks: a public university network; a private university network; a public and private university network.

We answer research question 3 by: first, analyzing overlap in the recruiting network of public universities (the public university network); second, analyzing overlap in the network of private universities (the private university network); and third, analyzing overlap amongst public and private universities (the public and private university network).

### Public university network

Tables \@ref(tab:pubu-count-matrix) and \@ref(tab:pubu-percent-matrix) provide a tabular representation of the extent to which public universities visited overlapping sets of private high schools. Table \@ref(tab:pubu-count-matrix) shows -- for each pair of public universities -- the number of high schools that were visited by both universities. For example, column 1 shows that of the 759 different private high schools that the University of Alabama visited, the University of South Carolina visited 284 of these schools and CU-Boulder visited 210 of these schools. Table \@ref(tab:pubu-percent-matrix) shows overlap in high schools visited in terms of percent rather than count. For example, column 2 shows that of the 396 different private high schools visited by the University of South Carolina, the University of Alabama visited 71.7% of these high schools and CU-Boulder visited 39.1% of these high schools.

INSERT TABLE \@ref(tab:pubu-count-matrix) AND TABLE \@ref(tab:pubu-percent-matrix) ABOUT HERE

One finding from Table \@ref(tab:pubu-percent-matrix) is that the University of Alabama is omnipresent in the set of private high schools visited by most other public research universities in the sample, particularly universities located in the South. For the remaining universities, Table \@ref(tab:pubu-percent-matrix) suggests several clusters of universities. For example, UMass-Amherst, SUNY Stony Brook, and Rutgers tend to visit many of the same high schools, as do the University of Arkansas and the University of Kansas.

Network graphs are useful for showing which universities have similar private high school recruiting networks. Figure \@ref(fig:plot-1mode-pubu) graphs the one-mode public university network. Whether graphing one-mode or two-mode networks, network graphs show vertices as circles and edges as lines connecting pairs of vertices. In Figure \@ref(fig:plot-1mode-pubu), each circle represents a public university and two universities share an edge -- represented by a solid line -- if both universities visit at least one high school in common. The width of the line between two universities indicates the number of high schools both universities visited, as identified in Table \@ref(tab:pubu-count-matrix). Conceptually, we think of vertices as repelling one another and edges as pulling vertices together [@RN4725]. Therefore, vertices that share few connections are located apart and those that share many connections are located close together. Vertices that are highly connected -- many direct and indirect connections to other nodes in the network -- are located near the center of the graph, while vertices that are relatively disconnected are located at the periphery of the graph.

INSERT FIGURE \@ref(fig:plot-1mode-pubu) ABOUT HERE

Figure \@ref(fig:plot-1mode-pubu) shows the University of Alabama, University of South Carolina, and CU-Boulder at the center of the recruiting network. The University of Nebraska-Lincoln, SUNY Stony Brook, and UC-Riverside are located on the periphery, and the remaining universities are in between. Universities that share many direct (i.e., visit many of the same schools) or indirect edges tend to be located close together, for example University of Kansas and University of Arkansas.

To better answer research question 3, we augment the network graphs by using cluster analysis methods -- referred to as "community detection" in the network analysis literature -- to identify groups of universities that share common network structure. Broadly speaking, the goal of community detection is to identify "subsets of vertices that demonstrate a ‘cohesiveness’ with respect to the underlying relational patterns" [@RN4725, p. 59], where a "cohesive subset" means a "subset of vertices that (i) are well connected among themselves, and at the same time (ii) are relatively well separated from the remaining vertices"  [@RN4725, p. 59].  In particular, we utilize that "cluster fast and greedy" approach developed by @RN4725, which is a variation of hierarchical cluster analysis that starts with a large number of groups and adds individual vertices to a group when doing so contributes to a cohesive subset.

In Figure \@ref(fig:plot-1mode-pubu), the shape of each university vertex is based on implementing this cluster fast and greedy approach with the specification that the algorithm should identify four groups.  The largest group consists of eight universities (U. Alabama, U. Arkansas, U. Kansas, Nebraska-Lincoln, U. Georgia, U. South Carolina, CU-Boulder, U. Cincinnati). With the exception of CU-Boulder, these universities were located in the South and the Midwest. Most of these universities made a relatively high share of visits in the South. The next largest cluster -- UMass Amherst, University of Pittsburgh, Rutgers, and SUNY Stony Brook -- consisted of universities that were mostly located in the Northeast and visited a large share of high schools in the Northeast. UC-Berkeley and UC-San Diego formed a two-university cluster and UC-Riverside -- which visited fewer private schools than any other public university -- was a cluster unto itself.

### Private university network

Table \@ref(tab:privu-percent-matrix) and Figure \@ref(fig:plot-1mode-privu) show overlap in private high schools visited by the selective private universities in our sample. The networks of private universities tend to have more overlap than the network of public research universities (above). This is demonstrated tabularly by the higher percentages in Table \@ref(tab:privu-percent-matrix) compared to Table \@ref(tab:pubu-percent-matrix) and graphically by the thicker lines connecting universities in Figure \@ref(fig:plot-1mode-privu) compared to Figure \@ref(fig:plot-1mode-pubu).

@@

INSERT FIGURE \@ref(fig:plot-1mode-privu) ABOUT HERE

Whereas applying cluster analysis to public universities (Figure \@ref(fig:plot-1mode-pubu)) yielded three clusters that contained multiple universities, the application of cluster analysis to private universities (Figure \@ref(fig:plot-1mode-privu)) yielded a single cluster that contained 11 universities and three clusters that contained one university each. As suggested above, a likely explanation for this difference is that the majority of private universities had substantial overlap in the sets of private high schools visited whereas public universities tended to have substantial overlap with some public universities but not others.

Visually, the most central universities in Figure \@ref(fig:plot-1mode-privu) tended to be universities that visited a large number of schools (number of unique private high schools visited in parentheses): Villanova (563), SMU (550), Notre Dame (625), Northwestern (377), Tulane (430), Boston College (339). The periphery of the network included Baylor (237), Stevens Institute (160), Marquette (331), and Case Western (228). Baylor visited a higher number of conservative Christian high schools than any other university (see Figure \@ref(fig:religion-pubu-privu)). It is notable that Marquette is in the periphery despite visiting a higher number of schools than more centrally located universities (e.g., Emory), suggesting that Marquette visits many schools that are visited by few other private universities in our sample.

The universities located close to one another in Figure \@ref(fig:plot-1mode-privu) tend to share religious affiliation and/or be located in the same geographic region. For example, the majority of Catholic universities are located close to one another and the majority of Midwest universities are located close to one another. Interestingly, two of the three Christian universities -- TCU and SMU -- are located much closer to the Catholic universities Notre Dame and Northwestern than they are to Baylor, suggesting that Baylor visits a different set of conservative Christian high schools than TCU and SMU.

### Public and private university network

Figure \@ref(fig:plot-1mode-u) shows overlap in private high schools visited amongst all 29 private and public universities in the sample.

The allocation of universities to clusters appears to be correlated with the characteristics -- university characteristics and characteristics of the schools they visit -- of geographic location, religious affiliation, academic reputation, and also institutional control. The cluster of 10 universities -- including the University of Alabama, TCU, SMU, Baylor, and the University of South Carolina -- largely consists of universities in the South, and includes both private Christian universities and public universities that visited a relatively large share of Christian high schools. Second, the cluster of 13 universities -- including Notre Dame, Northwestern, Tulane, CU-Boulder, UC-Berkeley -- largely consists of more highly ranked universities, universities located in the Midwest or West, and Catholic universities. Third, the five university cluster -- Rutgers, Stevens Institute of Technology, UMass Amherst, University of Pittsburgh, SUNY Stony Brook -- largely consists of public universities with strong ties to the Northeast.

@@

<!--

Tables \@ref(tab:u-cluster1-percent-matrix), \@ref(tab:u-cluster2-percent-matrix), and \@ref(tab:u-cluster3-percent-matrix) show overlap in the percent of private high schools 
INSERT TABLE \@ref(tab:u-cluster1-percent-matrix) TABLE \@ref(tab:u-cluster2-percent-matrix) AND TABLE \@ref(tab:u-cluster3-percent-matrix) ABOUT HERE
-->

Universities located near the center of Figure \@ref(fig:plot-1mode-u) do not all belong to the same cluster and, more broadly, universities located close to one another do not all belong to the same cluster. Universities located near the center of Figure \@ref(fig:plot-1mode-u) tended to be universities with the highest degree centrality (i.e., those that visited the most schools). The top nine universities in terms of degree centrality are (number of unique private high schools visited in parentheses): University of Alabama (759), Notre Dame (625), Villanova (563), SMU (550), TCU (435), Tulane (430), Northwestern (377), University of South Carolina (396), and CU-Boulder (362). Notably, while the majority of universities located near the center of Figure \@ref(fig:plot-1mode-u) are private, three are public.

Proximity of vertices to one another in Figure \@ref(fig:plot-1mode-u) appeared to be more connected to religious affiliation, geographic location, and academic reputation. Universities with a common religious affiliation (e.g., Notre Dame, Villanova, Marquette, Boston College) tend to be close to one another. However, the University of Alabama, which visited a high share of Catholic schools and Christian schools is simultaneously close to Catholic universities and to the Christian universities TCU and SMU. Similarly, the University of South Carolina, which visited a relatively high share of Christian schools, is close to TCU and SMU. CU-Boulder, Emory, and Northwestern -- three universities that differ from one another with respect to geographic location, religious affiliation, and institutional control -- are located close to one another in Figure \@ref(fig:plot-1mode-u). These three universities are similar to one another with respect to the geographic location (Figure \@ref(fig:region-pubu-privu)), religious affiliation (Figure \@ref(fig:religion-pubu-privu)), and academic reputation (Figure \@ref(fig:rank-pubu-privu)) of the private high schools they visit.

Finally, Table \@ref(tab:privu-percent-visited-by-pubu-matrix) shows, for each private university, the percent of schools that were also visited by each public university. This table shows the extent to which public universities were competing at private schools targeted by private universities. For example, of the 625 private schools visited by Notre Dame, the University of Alabama visited 52.5% of these schools. Similarly, CU-Boulder visited 46.4 of the 435 schools that TCU visited and 42.4% of the 377 schools that Northwestern visited. The University of Arkansas competed heavily at schools visited by Baylor; Rutgers and UMass Amherst competed heavily at schools visited by Stevens Institute of Technology; and UC-Berkeley competed at schools visited by Northwestern, Tufts, and Emory.

These analyses of the private and public university network provide the most direct assessment of research question 3. Considering the identification of clusters, centrality, and proximity of vertices, we find substantial overlap in the sets of private schools visited by public universities and private universities. This overlap does not suggest uniformity across public and private universities. Rather, there exist several groups of public and private universities -- which can be defined by cluster analysis or more loosely by proximity -- that have more in common with one another than they do with other universities of the same institutional control.

@@

Figures \@ref(fig:region-pubu-privu) through \@ref(fig:race-pubu-privhs-pubhs) address research question 3, "How do the characteristics of private high schools visited by public research universities compare to the characteristics of private high schools visited by selective private universities?"

Figure \@ref(fig:region-pubu-privu) shows the geographic region -- Northeast, Midwest, South, and West -- of visited private high schools, separately for public research universities (top panel) and selective private universities (bottom panel). Scholarship on the relationship between geographic proximity and student demand tends to find that nationally prestigious universities enjoy strong demand across the nation, while less prestigious universities rely primarily on demand from students located nearby [e.g., @RN2133]. The visit patterns for private universities are consistent with this finding; less prestigious universities tended to focus on private high schools in their home region, while more prestigious universities tended to visit many private high schools outside their home region.  Similarly, scholarship on student demand for out-of-state public universities finds that most public universities rely on demand from out-of-state students living in nearby states [@RN2536; @RN2535; @RN2546]. Most public research universities in our sample visited a disproportionate number of private high schools in their home geographic region. UC-Berkeley and CU Boulder -- universities with strong national brand recognition -- visited a large number of private high schools outside their home region.

@@

Figure \@ref(fig:rank-pubu-privu) shows the academic reputation -- defined by the Niche private high school ranking -- of visited private high schools. For the private universities in our sample, we see the predictable pattern that more highly ranked universities tend to visit a higher share of private high schools that are ranked in the top 200 nationally or are rated "A+" whereas less prestigious universities visited a higher share of private high schools rated "A" or "A- or below." Notre Dame University is a notable exception to this pattern. A potential explanation is that Notre Dame has developed a strategy of targeting the very top students at relatively less-prestigious Catholic high schools, believing that these students will choose Notre Dame because of its status as the preeminent Catholic University.

@@

A similar pattern emerged for the public research universities in our sample; more highly ranked universities tended to visit a higher share of highly rated private high schools than lower ranked universities. In comparison to the private universities in our sample, the public universities in our sample tended to visit a higher share of private high schools that were not highly rated. To a great extent, this finding is merely a consequence of our sampling strategy; we attempted to collect recruiting visit data from all public research universities but we only collected recruiting visit data from private universities ranked in the top 100 of U.S. News and World Report. That said, the finding that public research universities visited a large share of low-rated private high schools is consistent with the idea that many public research universities visit private high schools in search of full-pay students rather than academic superstars.

Figure \@ref(fig:enroll-pubu-privu) shows the 12th grade enrollment of visited private high schools. Compared to the universe of private high schools, the universities in our sample tended to visit schools that enrolled more 12th graders. Generally, the findings were pretty similar across public and private universities; for most universities, more than half of private school visits were to schools that enrolled at least 100 12th graders. However, public universities tended to visit a slightly larger share of high schools with fewer than 50 12th graders and private universities tended to visit a higher share of high schools with 12th grade enrollment of 50-100.

INSERT FIGURES \@ref(fig:enroll-pubu-privu) AND \@ref(fig:enroll-pubu-privhs-pubhs) ABOUT HERE

Figure \@ref(fig:enroll-pubu-privhs-pubhs) shows the 12th grade enrollment of visited private high schools as well as the 12th grade enrollment of visited public high schools for the public research universities in our sample. As seen, a large majority of public high school visits for all 15 universities were to schools with more than 150 12th graders enrolled (bottom panel). By contrast, most of the visited private schools tended to have fewer than 150 12th graders enrolled. Although this is somewhat to be expected, as public high schools tend to be larger in size than private schools (top rows of Figure \@ref(fig:enroll-pubu-privhs-pubhs)), there still appears to be a disproportionate amount of smaller private high schools visited compared to smaller public high schools. This suggests that the public universities in our sample were willing to take the time to visit private schools, even when they enroll far fewer 12th graders.

Figure \@ref(fig:race-pubu-privu) shows the percentage of 12th graders who identify as Black, Latinx, or Native at visited private high schools. Consistent with @RN4504, most private high schools enroll relatively few students from historically underrepresented race/ethnicity groups. For over 65%<!--0.6632124--> of schools in the universe of private high schools (top row of Figure \@ref(fig:race-pubu-privu)), Black/Latinx/Native students comprise less than 20% of enrolled students. Therefore, even if universities visited schools with racial composition on par with the universe of private high schools, most visited schools would have relatively few Black/Latinx/Native students enrolled. Figure \@ref(fig:race-pubu-privu) shows that for most universities in our sample, the set of visited private high schools enrolled *fewer* Black/Latinx/Native students than the universe of private high schools. For public universities, exceptions to this finding were UC-San Diego, UC-Riverside, and -- to a lesser extent -- SUNY Stony Brook. For private universities, Marquette was the only exception to this finding.

INSERT FIGURE \@ref(fig:race-pubu-privu) ABOUT HERE

@@

INSERT FIGURE \@ref(fig:race-pubu-privhs-pubhs) ABOUT HERE

Figure \@ref(fig:race-pubu-privhs-pubhs) compares the racial composition of private high schools visited by public research universities to the racial composition of public high schools visited by these universities. Looking at the universe of high schools (top rows of Figure \@ref(fig:race-pubu-privhs-pubhs)), we see that public high schools tend to be more racially diverse than private schools. This is reflected in the racial composition of visited schools as well, where the set of visited public high schools generally enroll more Black/Latinx/Native students than the visited private schools for the public research universities in our sample. Unfortunately though, a majority of these public universities tended to visit private high schools at a higher rate than public high schools, as seen earlier in Figure \@ref(fig:actual-proportional-pubu). This disproportionate emphasis on private school visits may shift the racial composition of public universities away from Black/Latinx/Native students.

>>>>

## RQ2: Overlapping network structure

Research question 2 asks, "To what extent do public research universities and selective private universities visit overlapping sets of private high schools?" Answering this question requires social network analysis techniques. As described in the methods section, we began by constructing a two-mode social network of visits to private high schools (mode 1) by universities (mode 2). Each university and each private high school is a "vertex." A visit from a university to a school is an "edge." Because large two-mode networks like our school-university network are difficult to visualize, two-mode networks are often analyzed as one-mode networks [@RN2457;@RN4663]. We transformed our two-mode school-university network into a one-mode university network in which each university is a vertex and two universities share an edge if both universities visited at least one high school in common. The *weight* of an edge indicates the number of high schools that both universities visited. For example, Appendix Table \@ref(tab:pubu-count-matrix) shows the matrix of weights for the set of public research universities. We created three separate one-mode university networks: a public university network; a private university network; a public and private university network.

We answer research question 2 by briefly describing tabular representations of overlap between the set of public research universities and overlap between the set of selective private universities. Next, we utilize tabular and graphical representations to describe ovelrap between the set of public and private universities in our sample.

Table \@ref(tab:pubu-percent-matrix) provide a tabular representation of the extent to which public universities visited overlapping sets of private high schools. For each public university (column), each cell of Table Table \@ref(tab:pubu-percent-matrix) shows number of visited high schools in common as a percent of the total number of high schools visited by the column university. This indicates the extent to which public universities were competing for students at the same set of private school. For example, column 2 shows that of the 396 different private high schools visited by the University of South Carolina, the University of Alabama visited 71.7% of these high schools and CU-Boulder visited 39.1% of these high schools. Appendix Table \@ref(tab:pubu-count-matrix) shows counts rather than percents.

INSERT TABLE \@ref(tab:pubu-percent-matrix) ABOUT HERE

One finding from Table \@ref(tab:pubu-percent-matrix) is that the University of Alabama is omnipresent in the set of private high schools visited by most other public research universities in the sample, particularly universities located in the South. For the remaining universities, Table \@ref(tab:pubu-percent-matrix) suggests several clusters of universities. For example, UMass-Amherst, SUNY Stony Brook, and Rutgers tend to visit many of the same high schools, as do the University of Arkansas and the University of Kansas. These descriptive findings suggest network structures based on geographic proximity.

Table \@ref(tab:privu-percent-matrix) show overlap in private high schools visited by the selective private universities in our sample. Appendix Table \@ref(tab:privu-count-matrix) shows counts. The networks of private universities appear to have more overlap than the network of public research universities. This is demonstrated tabularly by the higher percentages in Table \@ref(tab:privu-percent-matrix) compared to Table \@ref(tab:pubu-percent-matrix). Private universities that share a high percent of visits tend to have the same religious affiliation and/or be located in the same geographic region. For example, the Catholic universities -- Notre Dame, Villanova, Boston College, and Marquette -- seem to visit many private high schools in common.

@@

Next, we examine overlap in a network that contains the 15 public universities and the 14 private universities. Table \@ref(tab:privu-percent-visited-by-pubu-matrix) shows, for each private university, the percent of schools that were also visited by each public university. This table shows the extent to which public universities were competing at private schools targeted by private universities. For example, of the 625 private schools visited by Notre Dame, the University of Alabama visited 52.5% of these schools. Similarly, CU-Boulder visited 46.4 of the 435 schools that TCU visited and 42.4% of the 377 schools that Northwestern visited. The University of Arkansas competed heavily at schools visited by Baylor; Rutgers and UMass Amherst competed heavily at schools visited by Stevens Institute of Technology; and UC-Berkeley competed at schools visited by Northwestern, Tufts, and Emory. [CRYSTAL ADDS THE TABLE: FOR PUBLIC UNIVERSITIES, THE PRECENT OF HIGH SCHOOLS VISITED BY PRIVATE UNIVERSITIES]

Network graphs are useful for showing which universities have similar private high school recruiting networks. Figure \@ref(fig:plot-1mode-u) graphs the one-mode network that consists of all 29 public and private universities in the sample (Appendix Figure \@ref(fig:plot-1mode-pubu) graphs the one-mode public university network and Appendix Figure \@ref(fig:plot-1mode-privu) graph the one-mode private university network).

Whether graphing one-mode or two-mode networks, network graphs show vertices as circles and edges as lines connecting pairs of vertices. In Figure \@ref(fig:plot-1mode-u), each circle represents a public university and two universities share an edge -- represented by a solid line -- if both universities visit at least one high school in common. The width of the line between two universities indicates the *weight* of the edge, defined as number of high schools both universities visited. Network methods conceive of vertices as repelling one another [@RN4725], like the "north pole" of two magnets. By contrast, edges shared between two vertices pull the two vertices together, like the north pole of one magnet being attracted to the south pole of another magnet. Because of these laws of repulsion and attraction, vertices that share few connections are located far apart while vertices that share many connections are located close together. Additionally, vertices that are highly connected -- many direct and indirect connections to other nodes in the network -- are located near the center of the graph, while vertices that are relatively disconnected are located at the periphery of the graph.

@@

Universities located near the center of Figure \@ref(fig:plot-1mode-u) tended to be universities with the highest degree centrality (i.e., those that visited the most schools). The top nine universities in terms of degree centrality are (number of unique private high schools visited in parentheses): University of Alabama (759), Notre Dame (625), Villanova (563), SMU (550), TCU (435), Tulane (430), Northwestern (377), University of South Carolina (396), and CU-Boulder (362). Notably, while the majority of universities located near the center of Figure \@ref(fig:plot-1mode-u) are private, three are public. The most peripheral universities -- UC-Riverside,  SUNY Stony Brook, and University of Nebraska-Lincoln -- are all public.

Proximity of vertices to one another in Figure \@ref(fig:plot-1mode-u) appeared to be more connected to religious affiliation, geographic location, and academic reputation. Universities with a common religious affiliation (e.g., Notre Dame, Villanova, Marquette, Boston College) tend to be close to one another. However, the University of Alabama, which visited a high share of Catholic schools and Christian schools is simultaneously close to Catholic universities and to the Christian universities TCU and SMU. Similarly, the University of South Carolina, which visited a relatively high share of Christian schools, is close to TCU and SMU. CU-Boulder, Emory, and Northwestern -- three universities that differ from one another with respect to geographic location, religious affiliation, and institutional control -- are located close to one another in Figure \@ref(fig:plot-1mode-u). These three universities are similar to one another with respect to the geographic location (Figure \@ref(fig:region-pubu-privu)), religious affiliation (Figure \@ref(fig:religion-pubu-privu)), and academic reputation (Figure \@ref(fig:rank-pubu-privu)) of the private high schools they visit [HAVEN'T YET INTRODUCED THESE FIGURES].

To better answer research question 2, we augment the network graphs by using cluster analysis methods -- referred to as "community detection" in the network analysis literature -- to identify groups of universities that share common network structure. Broadly speaking, the goal of community detection is to identify "subsets of vertices that demonstrate a ‘cohesiveness’ with respect to the underlying relational patterns" [@RN4725, p. 59], where a "cohesive subset" means a "subset of vertices that (i) are well connected among themselves, and at the same time (ii) are relatively well separated from the remaining vertices"  [@RN4725, p. 59].  Therefore, universities in the same cluster have a relatively large number of direct and direct ties with other universities in the cluster and have relatively few direct and indirect ties with universities in other clusters. We utilize that "cluster fast and greedy" approach from @RN4725. This is a variation of hierarchical cluster analysis that starts with a large number of groups and adds individual vertices to a group when doing so contributes to a cohesive subset.

In Figure \@ref(fig:plot-1mode-u), universities are allocated to groups based on the specification that the algorithm should identify four groups (we also experimented with three, five, and six groups). Universities located near the center of Figure \@ref(fig:plot-1mode-u) do not all belong to the same cluster and, more broadly, universities located close to one another do not all belong to the same cluster. Starting with the most peripheral group, UC-Riverside is a group unto itself. UC-Riverside only visited 100 private schools and about 70 of these were located in CA. Another group --- Rutgers, Stevens Institute of Technology, UMass-Amherst, SUNY-Stony Brook, and Pitt -- largely consists of public universities with strong ties to the Northeast.

The remaining two groups are larger and include universities with greater network centrality.  Additionally, these groups both contain a mix of public and private universities. The cluster of 10 universities -- including the University of Alabama, TCU, SMU, Baylor, University of Kansas, and the University of South Carolina -- largely consists of universities in the South, and includes both private Christian universities and public universities. While the federal government defines Nebraska and Kansas as part of the Midwest, Figure \@ref(fig:plot-1mode-u) suggests that the visit patterns of the University of Kansas and the University of Nebraska share more in common with Southern universities than they do with other universities in our sample. Finally, the largest and most centrally located cluster contains 13 universities, including Notre Dame, Northwestern, Tulane, CU-Boulder, UC-Berkeley. Compared to the overall sample, this group tends to be. This group largely consists of highly ranked universities, universities located in the Midwest or West, and Catholic universities.

The allocation of universities to clusters appears to be correlated with the characteristics -- university characteristics and characteristics of the schools they visit -- of geographic location, religious affiliation, academic reputation, and also institutional control. Considering the identification of clusters, centrality, and proximity of vertices, we find substantial overlap in the sets of private schools visited by public universities and private universities. This overlap does not suggest uniformity across public and private universities. Rather, there exist several groups of public and private universities -- which can be defined by cluster analysis or more loosely by proximity -- that have more in common with one another than they do with other universities of the same institutional control.

@@

Figures \@ref(fig:region-pubu-privu) through \@ref(fig:race-pubu-privhs-pubhs) address research question 3, "How do the characteristics of private high schools visited by public research universities compare to the characteristics of private high schools visited by selective private universities?" Additionally, Appendix Figures \@ref(fig:enroll-pubu-privu) and \@ref(fig:enroll-pubu-privhs-pubhs) show the 12th grade enrollment size of visited high schools.

Figure \@ref(fig:region-pubu-privu) shows the broad geographic region -- Northeast, Midwest, South, and West -- of visited private high schools, separately for public research universities (top panel) and selective private universities (bottom panel). Scholarship on the relationship between geographic proximity and student demand tends to find that nationally prestigious universities enjoy strong demand across the nation, while less prestigious universities rely primarily on demand from students located nearby [e.g., @RN2133]. The visit patterns for private universities are consistent with this finding; less prestigious universities tended to focus on private high schools in their home region, while more prestigious universities tended to visit many private high schools outside their home region.  Similarly, scholarship on student demand for out-of-state public universities finds that most public universities rely on demand from out-of-state students living in nearby states [@RN2536; @RN2535; @RN2546]. Most public research universities in our sample visited a disproportionate number of private high schools in their home geographic region. UC-Berkeley and CU Boulder -- universities with strong national brand recognition -- visited a large number of private high schools outside their home region.

@@

Figure \@ref(fig:rank-pubu-privu) shows the academic reputation -- defined by the Niche private high school ranking -- of visited private high schools. For the private universities in our sample, we see the predictable pattern that more highly ranked universities tend to visit a higher share of private high schools that are ranked in the top 200 nationally or are rated "A+" whereas less prestigious universities visited a higher share of private high schools rated "A" or "A- or below." Notre Dame University is a notable exception to this pattern.^[A potential explanation is that Notre Dame has developed a strategy of targeting the very top students at relatively less-prestigious Catholic high schools, believing that these students will choose Notre Dame because of its status as the preeminent Catholic University.]

@@

A similar pattern emerged for the public research universities in our sample; more highly ranked universities tended to visit a higher share of highly rated private high schools than lower ranked universities. In comparison to the private universities in our sample, the public universities in our sample tended to visit a higher share of private high schools that were not highly rated. To a great extent, this finding is merely a consequence of our sampling strategy; we attempted to collect recruiting visit data from all public research universities but we only collected recruiting visit data from private universities ranked in the top 100 of U.S. News and World Report. That said, the finding that public research universities visited a large share of low-rated private high schools is consistent with the idea that many public research universities visit private high schools in search of full-pay students rather than high-achieving students.

Figures \@ref(fig:race-pubu-privhs-pubhs) and \@ref(fig:race-privu-privhs-pubhs) shows the percentage of 12th graders who identify as Black, Latinx, or Native at visited high schools. Consistent with @RN4504, most private high schools enroll relatively few students from historically underrepresented race/ethnicity groups. The top row in the top panel of Figure \@ref(fig:race-pubu-privhs-pubhs) shows that the sum of Black, Latinx, and Native students comprise less than 20% of enrolled students in over over 65%<!--0.6632124--> of schools in private high school universe. At public high schools, Black/Latinx/Native students comprise less than 20% of enrolled students at less than 40% [CRYSTAL CHECK NUMBER] of all public schools.

Figure \@ref(fig:race-pubu-privhs-pubhs) focuses on public universities, showing the racial composition of visited private high schools (top panel) to the racial composition of visited public high schools (bottom panel). START HERE TUESDAY. CONSISTENT WITH DIFFERENCES IN RACIAL COMPOSITION OF THE UNIVERSE OF PUBLIC VS. THE UNIVERSE OF PRIVATE, VISITED PRIVATE SCHOOLS TENDED TO ENROLL A MUCH LOWER SHARE OF BLACK/LATINX THAN VISITED PUBLIC SCHOOLS.

This is reflected in the racial composition of visited schools as well, where the set of visited public high schools generally enroll more Black/Latinx/Native students than the visited private schools for the public research universities in our sample. Unfortunately though, a majority of these public universities tended to visit private high schools at a higher rate than public high schools, as seen earlier in Figure \@ref(fig:actual-proportional-pubu). This disproportionate emphasis on private school visits may shift the racial composition of public universities away from Black/Latinx/Native students.

Therefore, even if universities visited schools with racial composition on par with the universe of private high schools, most visited schools would have relatively few Black/Latinx/Native students enrolled.

Figure \@ref(fig:race-pubu-privu) shows that for most universities in our sample, the set of visited private high schools enrolled *fewer* Black/Latinx/Native students than the universe of private high schools. For public universities, exceptions to this finding were UC-San Diego, UC-Riverside, and -- to a lesser extent -- SUNY Stony Brook. For private universities, Marquette was the only exception to this finding.

INSERT FIGURES \@ref(fig:race-pubu-privhs-pubhs) AND \@ref(fig:race-privu-privhs-pubhs) ABOUT HERE

@@

ADD ADDITIONAL SUBGROUP ANALYSES WITH CRYSTAL...?

- look at racial composition of visited southern schools (public and private)
- look at religious affiliation of visited southern schools?
- look at racial composition of schools visited by some subset of universities?
-
